Distributed multi-objective evolutionary optimization using island-based selective operator application

### 2018-06-04 hpmoon-asoco.tex
Changed Adaptive to Automatic. Closes #24

In order to prove this claim, our new adaptive method has been
compared against different overlapping island schemes: a
baseline version of a distributed NSGA-II \citep{Deb00NSGAII}, a
SLA version with disjoint sections and a SLA
fixed-size overlapped version. 
Several benchmark problems and different large number of islands
have been used. Results show that this new technique can improve
different quality indicators in the same amount of time when the
problem size is large.

@@

In our previous work \citep{Garcia16hpmoon} we used some of
the aforementioned ideas to compare two different dMOEAs. The first
one divided the chromosome into $P$ sections, being $P$ the number of
islands. Every island $p$ only performed the mutation and crossover in
that part ($p_{th}$) of the chromosome (the selective operator application), while the fitness was
calculated using the whole individual. After a certain number of
generations, individuals were migrated randomly to other islands. The
performance metrics were calculated at the end of the run. The second
method, selective operator application with overlapping islands, used the $p_{th-1}$ and
$p_{th+1}$ sections of each chromosome, besides the $p_{th}$
part. During the same amount of time both methods obtained better
results than a baseline algorithm that dealt with the whole chromosome
in each island for crossover and mutation. We discovered that the
performance using one or another method depends on the number of
sections of the individuals and the number of islands used. This
motivated us to find a new adaptive method to select this number of
sections of the chromosome to use, depending on the number of
islands. Besides, previous experiments were performed on a single
processor synchronous island model with a limited number of islands
(8, 32 and 128). In this paper,  8, 16, 32, 64 and 128 islands have
been used, and on this occasion, the experiments have been run on a parallel
cluster. Therefore, at the same time we are proposing an adaptive
method to divide the individual search space, and also validating the
previous approach.

@@

\subsubsection{Adaptive SLA with overlapping islands (A)} 
As in the previous method, the sections to address by the operators are overlapped, but
instead of using one extra section on each side of the $p$ section ($p+1$
and $p-1$), it uses $c$ fragments on each side ($p+c$ and $p-c$),
being $c$ a value depending on the number of islands.
% That's not adaptive. It's simply variable - JJ FERGU: it depends on the number of islands. What term do you suggest?

As a first
approach to automatically calculate this value, the results shown in
\citep{Garcia16hpmoon} have been used as a base to obtain $c$. In that
work, the overlapped method ($c=1$) obtained better results when the
number of islands were bigger than 8. On the contrary, when the number of
islands is small, no overlapping is necessary. So, we have used this
knowledge to create the formula $c=round(0.2*P-1)/2$ to calculate the
extra sections to overlap. Therefore, when $P=8$ then $c=0$
(equivalent to D), when $P=16$ then $c=1$ (equivalent to O), and so
on. \textcolor{blue}{Dividing individuals by blocks of equal size, instead a percentage, allow the future integration of our method with other recombination mechanisms, such as the ones presented in \cite{Dorronsoro13superlinear} or \cite{Kimovski15Parallel}}.
Also, as mentioned, this is only a first approach to obtain $c$, and future work will study new methods to calculate it. Figure \ref{fig:adaptive} explains this method.

\begin{figure*}[h!tb]
\centering
\includegraphics[width=12cm]{islandAdaptive.jpg}
\caption{Adaptive SLA with overlapping islands (A): every island $p$ modifies the  $p+c$,
  $p_{th}$ and $p-c$  components (in grey) of the individuals. $c$ is calculated depending on the number of islands. In this case, $c=2$.}
  \label{fig:adaptive}
\end{figure*}
% --------------------------------------------------------------

@@

\begin{algorithmic}
\STATE \textit{//In each island $i$}
\STATE population $\gets$ initialisePopulation()
\STATE \textit{//Select which part modify in this island $i$}
\STATE section $\gets$ getComponentsToModify($i$)
\IF {type = Baseline}
	\STATE $c=P/2$
\ELSIF {type = Disjoint}
	\STATE $c=0$
\ELSIF {type = Overlapping}
	\STATE $c=1$
\ELSIF {type = Adaptive}
	\STATE $c=round(0.2*P-1)/2$
\ENDIF

@@

\end{tabular}
}
\caption{Average quality metrics obtained after 30 runs per configuration, for the 4 methods compared: baseline (B), disjoint (D), overlapped (O) and adaptive overlapped (A), using a chromosome length of 512 dimensions. Acronyms next to values indicate that there is not significant difference with respect to that method for that value. Best values are marked in bold. Equiv-$X$ implies that value is the same as executing $X$, since both configurations would be the same.}
\label{tab:results512}
\end{table*}

@@

\end{tabular}
}
\caption{Average quality metrics obtained after 30 runs per configuration, for the 4 methods compared: baseline (B), disjoint (D), overlapped (O) and adaptive overlapped (A), using a chromosome length of 2048 dimensions. Acronyms next to values indicate that there is not significant difference with respect to that method for that value. Best values are marked in bold. Equiv-$X$ implies that value is the same as executing $X$, since both configurations would be the same.}
\label{tab:results2048}
\end{table*}

@@

\#Island	&	B		&	D		&	O			&	A			&	B		&	D			&	O			&	A				\\ \hline
\multicolumn{9}{|c|}{ZDT1}																				\\ \hline
8	&	182.700		&	128.500		&	158.167			&	Equiv-D			&	397.367		&	776.967			&	603.667			&	Equiv-D				\\
16	&	132.567		&	43.633		&	76.167			&	Equiv-O			&	567.933		&	908.467			&	838.000			&	Equiv-O				\\
32	&	102.200		&	37.533		&	34.133	D		&	52.033			&	731.067		&	959.067			&	938.167			&	917.000				\\
64	&	81.100		&	46.267		&	26.300			&	30.167		O	&	846.733		&	978.733			&	972.567			&	927.133	B			\\
128	&	77.167		&	57.767		&	35.767			&	29.167		O	&	923.133		&	984.767			&	984.233	D		&	967.933				\\ \hline
\multicolumn{9}{|c|}{ZDT2}																															\\ \hline
8	&	122.700		&	91.633		&	108.833	BD		&	Equiv-D			&	398.167		&	772.000			&	605.100			&	Equiv-D				\\
16	&	94.867		&	28.033		&	62.467			&	Equiv-O			&	574.567		&	908.433			&	840.900			&	Equiv-O				\\
32	&	56.633		&	11.433		&	19.467	D		&	31.567			&	735.867		&	960.167			&	939.200			&	904.900				\\
64	&	33.467		&	10.267		&	9.367	D		&	19.067			&	849.900		&	978.600			&	972.167			&	929.800	B			\\
128	&	22.833		&	12.633		&	7.800			&	13.967		D	&	929.433		&	984.567			&	984.233	D		&	971.700				\\ \hline
\multicolumn{9}{|c|}{ZDT3}																															\\ \hline
8	&	212.900		&	129.233		&	171.367			&	Equiv-D			&	398.633		&	776.000			&	603.333			&	Equiv-D				\\
16	&	176.133		&	43.567		&	90.100			&	Equiv-O			&	573.033		&	908.667			&	838.000			&	Equiv-O				\\
32	&	121.500		&	42.800		&	33.200			&	56.900			&	733.233		&	959.000			&	937.967			&	918.833				\\
64	&	96.500		&	54.267		&	34.900			&	37.800		O	&	849.200		&	977.167			&	972.200			&	933.167				\\
128	&	76.700		&	62.200		&	41.767			&	33.867		D	&	924.000		&	984.733			&	984.367	D		&	966.433				\\ \hline
\multicolumn{9}{|c|}{ZDT6}																															\\ \hline
8	&	42.233		&	28.633		&	35.967	BD		&	Equiv-D			&	396.900		&	773.400			&	603.533			&	Equiv-D				\\
16	&	43.733		&	7.733		&	18.967			&	Equiv-O			&	568.200		&	908.900			&	836.033			&	Equiv-O				\\
32	&	35.533		&	8.033		&	8.633	D		&	16.233		DO	&	735.700		&	958.900			&	938.600			&	926.167				\\
64	&	22.600		&	10.700		&	7.500	D		&	7.533		DO	&	850.267		&	978.067			&	972.467			&	940.267				\\
128	&	18.133		&	15.000	B	&	9.067			&	11.367		D	&	927.233		&	984.667			&	983.933	D		&	969.433				\\ \hline
\end{tabular}
}
\caption{Average number of generations and average number of solutions of the front, obtained after 30 runs per configuration, for the 4 methods compared: baseline (B), disjoint (D), overlapped (O) and adaptive overlapped (A), using a chromosome length of 512 dimensions. Acronyms next to values indicate that there is not significant difference with respect to that method for that value. Equiv-$X$ implies that value is the same as executing $X$, since both configurations would be the same.}
\label{tab:sols512}
\end{table*}

\begin{table*}
\centering
\resizebox{12cm}{!}{
\begin{tabular}{|c||c|c|c|c||c|c|c|c||}
\hline
	&	\multicolumn{4}{|c|}{Average solutions per front}													&	\multicolumn{4}{|c|}{Generations}															\\ \hline
%\multicolumn{13}{|c|}{2048 dimensions}																															\\ \hline
\#Island	&	B		&	D		&	O			&	A			&	B		&	D			&	O			&	A				\\ \hline
\multicolumn{9}{|c|}{ZDT1}																					\\ \hline
8	&	137.733		&	47.167		&	119.933	B		&	Equiv-D			&	175.067		&	225.667			&	207.933			&	Equiv-D				\\
16	&	92.633		&	38.233		&	47.533	D		&	Equiv-O			&	204.867		&	238.533			&	232.900			&	Equiv-O				\\
32	&	56.167		&	41.167		&	28.667			&	32.567	O		&	225.433		&	243.833			&	242.533			&	242.000	O			\\
64	&	50.900		&	49.667	B	&	35.600			&	25.367			&	236.500		&	245.967			&	245.700			&	244.033	D			\\
128	&	45.767		&	64.933		&	44.167	B		&	31.867			&	242.800		&	247.000			&	247.000	D		&	245.733				\\ \hline
\multicolumn{9}{|c|}{ZDT2}																															\\ \hline
8	&	64.267		&	22.733		&	56.867	B		&	Equiv-D			&	175.633		&	226.000			&	208.100			&	Equiv-D				\\
16	&	52.300		&	10.733		&	20.067	D		&	Equiv-O			&	205.033		&	238.867			&	233.100			&	Equiv-O				\\
32	&	34.400		&	10.367		&	10.433	D		&	17.900	O		&	225.600		&	244.267			&	242.633			&	242.000	O			\\
64	&	24.600		&	10.900		&	9.267	D		&	12.867	DO		&	236.700		&	246.000			&	245.767			&	244.033	D			\\
128	&	18.667		&	17.833	B	&	10.367			&	12.933	O		&	243.367		&	247.000			&	247.000	d		&	245.833				\\ \hline
\multicolumn{9}{|c|}{ZDT3}																															\\ \hline
8	&	163.767		&	83.367		&	119.467			&	Equiv-D			&	176.533		&	226.400			&	207.700			&	Equiv-D				\\
16	&	108.467		&	49.433		&	47.700	D		&	Equiv-O			&	205.333		&	238.467			&	232.933			&	Equiv-O				\\
32	&	72.133		&	44.333		&	31.833			&	36.533	O		&	225.100		&	243.900			&	242.433			&	242.000	O			\\
64	&	54.867		&	57.600	B	&	40.633			&	29.300			&	236.333		&	245.933			&	245.733			&	244.000	D			\\
128	&	50.600		&	72.133		&	49.033	B		&	34.933			&	243.200		&	247.000			&	247.000	D		&	245.833				\\ \hline
\multicolumn{9}{|c|}{ZDT6}																															\\ \hline
8	&	22.133		&	13.833		&	14.433	D		&	Equiv-D			&	175.733		&	226.100			&	207.767			&	Equiv-D				\\
16	&	20.767		&	15.867		&	13.300	D		&	Equiv-O			&	205.033		&	238.433			&	232.933			&	Equiv-O				\\
32	&	22.600		&	10.700		&	10.033	D		&	11.967	DO		&	225.833		&	243.667			&	242.500			&	242.000	O			\\
64	&	19.033		&	11.267		&	10.533	D		&	10.567	DO		&	236.433		&	245.900			&	245.833			&	244.000	D			\\
128	&	15.800		&	24.500		&	11.600			&	12.233	DO		&	243.800		&	247.000			&	247.000	D		&	245.533				\\ \hline
\end{tabular}
}
\caption{Average number of generations and average number of solutions per front, obtained after 30 runs per configuration, for the 4 methods compared: baseline (B), disjoint (D), overlapped (O) and adaptive overlapped (A), using a chromosome length of 2048 dimensions. Acronyms next to values indicate that there is not significant difference with respect to that method for that value. Equiv-$X$ implies that value is the same as executing $X$, since both configurations would be the same.}
\label{tab:sols2048}
\end{table*}

@@

However, with 2048 dimensions, results show that dividing the chromosome produces an improvement in all the quality indicators using the adaptive version (Table \ref{tab:results2048}), even improving the overlapping and disjoint methods. Therefore, there exist some kind of limit point in chromosome length where one method will be preferable to another, besides depending on the number of islands and population size.

\begin{figure}
\centering
\includegraphics[width=12cm]{plot_zdt1_512.png}
\caption{\textcolor{blue}{PFs obtained in one arbitrary run for the ZDT1 problem with a chromosome size ($L$) of 512 for each combination of number of islands ($P$) and method: baseline ($B$), disjoint ($D$), overlapped ($O$) and adaptive overlapped ($A$). Results for 8 and 16 islands of $A$ are equivalent of running $D$ and $O$ respectively. }}
\label{fig:plot_zdt1_512}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=12cm]{plot_zdt2_512.png}
\caption{\textcolor{blue}{PFs obtained in one arbitrary run for the ZDT2 problem with a chromosome size ($L$) of 512 for each combination of number of islands ($P$) and method: baseline ($B$), disjoint ($D$), overlapped ($O$) and adaptive overlapped ($A$). Results for 8 and 16 islands of $A$ are equivalent of running $D$ and $O$ respectively. }}
\label{fig:plot_zdt2_512}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=12cm]{plot_zdt3_512.png}
\caption{\textcolor{blue}{PFs obtained in one arbitrary run for the ZDT3 problem with a chromosome size ($L$) of 512 for each combination of number of islands ($P$) and method: baseline ($B$), disjoint ($D$), overlapped ($O$) and adaptive overlapped ($A$). Results for 8 and 16 islands of $A$ are equivalent of running $D$ and $O$ respectively. }}
\label{fig:plot_zdt3_512}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=12cm]{plot_zdt6_512.png}
\caption{\textcolor{blue}{PFs obtained in one arbitrary run for the ZDT6 problem with a chromosome size ($L$) of 512 for each combination of number of islands ($P$) and method: baseline ($B$), disjoint ($D$), overlapped ($O$) and adaptive overlapped ($A$). Results for 8 and 16 islands of $A$ are equivalent of running $D$ and $O$ respectively. }}
\label{fig:plot_zdt6_512}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=12cm]{plot_zdt1_2048.png}
\caption{\textcolor{blue}{PFs obtained in one arbitrary run for the ZDT1 problem with a chromosome size ($L$) of 2048 for each combination of number of islands ($P$) and method: baseline ($B$), disjoint ($D$), overlapped ($O$) and adaptive overlapped ($A$). Results for 8 and 16 islands of $A$ are equivalent of running $D$ and $O$ respectively. }}
\label{fig:plot_zdt1_2048}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=12cm]{plot_zdt2_2048.png}
\caption{\textcolor{blue}{PFs obtained in one arbitrary run for the ZDT2 problem with a chromosome size ($L$) of 2048 for each combination of number of islands ($P$) and method: baseline ($B$), disjoint ($D$), overlapped ($O$) and adaptive overlapped ($A$). Results for 8 and 16 islands of $A$ are equivalent of running $D$ and $O$ respectively. }}
\label{fig:plot_zdt2_2048}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=12cm]{plot_zdt3_2048.png}
\caption{\textcolor{blue}{PFs obtained in one arbitrary run for the ZDT3 problem with a chromosome size ($L$) of 2048 for each combination of number of islands ($P$) and method: baseline ($B$), disjoint ($D$), overlapped ($O$) and adaptive overlapped ($A$). Results for 8 and 16 islands of $A$ are equivalent of running $D$ and $O$ respectively. }}
\label{fig:plot_zdt3_2048}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=12cm]{plot_zdt6_2048.png}
\caption{\textcolor{blue}{PFs obtained in one arbitrary run for the ZDT6 problem with a chromosome size ($L$) of 2048 for each combination of number of islands ($P$) and method: baseline ($B$), disjoint ($D$), overlapped ($O$) and adaptive overlapped ($A$). Results for 8 and 16 islands of $A$ are equivalent of running $D$ and $O$ respectively. }}
\label{fig:plot_zdt6_2048}
\end{figure}

>>>>

In order to prove this claim, our new automatic method has been
compared against different overlapping island schemes: a
baseline version of a distributed NSGA-II \citep{Deb00NSGAII}, a
SLA version with disjoint sections and a SLA
fixed-size overlapped version. 
Several benchmark problems and different large number of islands
have been used. Results show that this new technique can improve
different quality indicators in the same amount of time when the
problem size is large.

@@

In our previous work \citep{Garcia16hpmoon} we used some of
the aforementioned ideas to compare two different dMOEAs. The first
one divided the chromosome into $P$ sections, being $P$ the number of
islands. Every island $p$ only performed the mutation and crossover in
that part ($p_{th}$) of the chromosome (the selective operator application), while the fitness was
calculated using the whole individual. After a certain number of
generations, individuals were migrated randomly to other islands. The
performance metrics were calculated at the end of the run. The second
method, selective operator application with overlapping islands, used the $p_{th-1}$ and
$p_{th+1}$ sections of each chromosome, besides the $p_{th}$
part. During the same amount of time both methods obtained better
results than a baseline algorithm that dealt with the whole chromosome
in each island for crossover and mutation. We discovered that the
performance using one or another method depends on the number of
sections of the individuals and the number of islands used. This
motivated us to find a new automatic method to select this number of
sections of the chromosome to use, depending on the number of
islands. Besides, previous experiments were performed on a single
processor synchronous island model with a limited number of islands
(8, 32 and 128). In this paper,  8, 16, 32, 64 and 128 islands have
been used, and on this occasion, the experiments have been run on a parallel
cluster. Therefore, at the same time we are proposing an automatic
method to divide the individual search space, and also validating the
previous approach.

@@

\subsubsection{\textcolor{blue}{Automatic} SLA with overlapping islands (A)} 
As in the previous method, the sections to address by the operators are overlapped, but
instead of using one extra section on each side of the $p$ section ($p+1$
and $p-1$), it uses $c$ fragments on each side ($p+c$ and $p-c$),
being $c$ a value depending on the number of islands.
% That's not adaptive. It's simply variable - JJ FERGU: it depends on the number of islands. What term do you suggest?
% FERGU: changed to Automatic
As a first
approach to automatically calculate this value, the results shown in
\citep{Garcia16hpmoon} have been used as a base to obtain $c$. In that
work, the overlapped method ($c=1$) obtained better results when the
number of islands were bigger than 8. On the contrary, when the number of
islands is small, no overlapping is necessary. So, we have used this
knowledge to create the formula $c=round(0.2*P-1)/2$ to calculate the
extra sections to overlap. Therefore, when $P=8$ then $c=0$
(equivalent to D), when $P=16$ then $c=1$ (equivalent to O), and so
on. \textcolor{blue}{Dividing individuals by blocks of equal size, instead a percentage, allow the future integration of our method with other recombination mechanisms, such as the ones presented in \cite{Dorronsoro13superlinear} or \cite{Kimovski15Parallel}}.
Also, as mentioned, this is only a first approach to obtain $c$, and future work will study new methods to calculate it. Figure \ref{fig:adaptive} explains this method.

\begin{figure*}[h!tb]
\centering
\includegraphics[width=12cm]{islandAdaptive.jpg}
\caption{\textcolor{blue}{Automatic} SLA with overlapping islands (A): every island $p$ modifies the  $p+c$,
  $p_{th}$ and $p-c$  components (in grey) of the individuals. $c$ is calculated depending on the number of islands. In this case, $c=2$.}
  \label{fig:adaptive}
\end{figure*}
% --------------------------------------------------------------

@@

\begin{algorithmic}
\STATE \textit{//In each island $i$}
\STATE population $\gets$ initialisePopulation()
\STATE \textit{//Select which part modify in this island $i$}
\STATE section $\gets$ getComponentsToModify($i$)
\IF {type = Baseline}
	\STATE $c=P/2$
\ELSIF {type = Disjoint}
	\STATE $c=0$
\ELSIF {type = Overlapping}
	\STATE $c=1$
\ELSIF {type = Automatic}
	\STATE $c=round(0.2*P-1)/2$
\ENDIF

@@

\end{tabular}
}
\caption{Average quality metrics obtained after 30 runs per configuration, for the 4 methods compared: baseline (B), disjoint (D), overlapped (O) and \textcolor{blue}{automatic} overlapped (A), using a chromosome length of 512 dimensions. Acronyms next to values indicate that there is not significant difference with respect to that method for that value. Best values are marked in bold. Equiv-$X$ implies that value is the same as executing $X$, since both configurations would be the same.}
\label{tab:results512}
\end{table*}

@@

\end{tabular}
}
\caption{Average quality metrics obtained after 30 runs per configuration, for the 4 methods compared: baseline (B), disjoint (D), overlapped (O) and \textcolor{blue}{automatic} overlapped (A), using a chromosome length of 2048 dimensions. Acronyms next to values indicate that there is not significant difference with respect to that method for that value. Best values are marked in bold. Equiv-$X$ implies that value is the same as executing $X$, since both configurations would be the same.}
\label{tab:results2048}
\end{table*}

@@

\#Island	&	B		&	D		&	O			&	A			&	B		&	D			&	O			&	A				\\ \hline
\multicolumn{9}{|c|}{ZDT1}																				\\ \hline
8	&	182.700		&	128.500		&	158.167			&	Equiv-D			&	397.367		&	776.967			&	603.667			&	Equiv-D				\\
16	&	132.567		&	43.633		&	76.167			&	Equiv-O			&	567.933		&	908.467			&	838.000			&	Equiv-O				\\
32	&	102.200		&	37.533		&	34.133	D		&	52.033			&	731.067		&	959.067			&	938.167			&	917.000				\\
64	&	81.100		&	46.267		&	26.300			&	30.167		O	&	846.733		&	978.733			&	972.567			&	927.133	B			\\
128	&	77.167		&	57.767		&	35.767			&	29.167		O	&	923.133		&	984.767			&	984.233	D		&	967.933				\\ \hline
\multicolumn{9}{|c|}{ZDT2}																															\\ \hline
8	&	122.700		&	91.633		&	108.833	BD		&	Equiv-D			&	398.167		&	772.000			&	605.100			&	Equiv-D				\\
16	&	94.867		&	28.033		&	62.467			&	Equiv-O			&	574.567		&	908.433			&	840.900			&	Equiv-O				\\
32	&	56.633		&	11.433		&	19.467	D		&	31.567			&	735.867		&	960.167			&	939.200			&	904.900				\\
64	&	33.467		&	10.267		&	9.367	D		&	19.067			&	849.900		&	978.600			&	972.167			&	929.800	B			\\
128	&	22.833		&	12.633		&	7.800			&	13.967		D	&	929.433		&	984.567			&	984.233	D		&	971.700				\\ \hline
\multicolumn{9}{|c|}{ZDT3}																															\\ \hline
8	&	212.900		&	129.233		&	171.367			&	Equiv-D			&	398.633		&	776.000			&	603.333			&	Equiv-D				\\
16	&	176.133		&	43.567		&	90.100			&	Equiv-O			&	573.033		&	908.667			&	838.000			&	Equiv-O				\\
32	&	121.500		&	42.800		&	33.200			&	56.900			&	733.233		&	959.000			&	937.967			&	918.833				\\
64	&	96.500		&	54.267		&	34.900			&	37.800		O	&	849.200		&	977.167			&	972.200			&	933.167				\\
128	&	76.700		&	62.200		&	41.767			&	33.867		D	&	924.000		&	984.733			&	984.367	D		&	966.433				\\ \hline
\multicolumn{9}{|c|}{ZDT6}																															\\ \hline
8	&	42.233		&	28.633		&	35.967	BD		&	Equiv-D			&	396.900		&	773.400			&	603.533			&	Equiv-D				\\
16	&	43.733		&	7.733		&	18.967			&	Equiv-O			&	568.200		&	908.900			&	836.033			&	Equiv-O				\\
32	&	35.533		&	8.033		&	8.633	D		&	16.233		DO	&	735.700		&	958.900			&	938.600			&	926.167				\\
64	&	22.600		&	10.700		&	7.500	D		&	7.533		DO	&	850.267		&	978.067			&	972.467			&	940.267				\\
128	&	18.133		&	15.000	B	&	9.067			&	11.367		D	&	927.233		&	984.667			&	983.933	D		&	969.433				\\ \hline
\end{tabular}
}
\caption{Average number of generations and average number of solutions of the front, obtained after 30 runs per configuration, for the 4 methods compared: baseline (B), disjoint (D), overlapped (O) and \textcolor{blue}{automatic} overlapped (A), using a chromosome length of 512 dimensions. Acronyms next to values indicate that there is not significant difference with respect to that method for that value. Equiv-$X$ implies that value is the same as executing $X$, since both configurations would be the same.}
\label{tab:sols512}
\end{table*}

\begin{table*}
\centering
\resizebox{12cm}{!}{
\begin{tabular}{|c||c|c|c|c||c|c|c|c||}
\hline
	&	\multicolumn{4}{|c|}{Average solutions per front}													&	\multicolumn{4}{|c|}{Generations}															\\ \hline
%\multicolumn{13}{|c|}{2048 dimensions}																															\\ \hline
\#Island	&	B		&	D		&	O			&	A			&	B		&	D			&	O			&	A				\\ \hline
\multicolumn{9}{|c|}{ZDT1}																					\\ \hline
8	&	137.733		&	47.167		&	119.933	B		&	Equiv-D			&	175.067		&	225.667			&	207.933			&	Equiv-D				\\
16	&	92.633		&	38.233		&	47.533	D		&	Equiv-O			&	204.867		&	238.533			&	232.900			&	Equiv-O				\\
32	&	56.167		&	41.167		&	28.667			&	32.567	O		&	225.433		&	243.833			&	242.533			&	242.000	O			\\
64	&	50.900		&	49.667	B	&	35.600			&	25.367			&	236.500		&	245.967			&	245.700			&	244.033	D			\\
128	&	45.767		&	64.933		&	44.167	B		&	31.867			&	242.800		&	247.000			&	247.000	D		&	245.733				\\ \hline
\multicolumn{9}{|c|}{ZDT2}																															\\ \hline
8	&	64.267		&	22.733		&	56.867	B		&	Equiv-D			&	175.633		&	226.000			&	208.100			&	Equiv-D				\\
16	&	52.300		&	10.733		&	20.067	D		&	Equiv-O			&	205.033		&	238.867			&	233.100			&	Equiv-O				\\
32	&	34.400		&	10.367		&	10.433	D		&	17.900	O		&	225.600		&	244.267			&	242.633			&	242.000	O			\\
64	&	24.600		&	10.900		&	9.267	D		&	12.867	DO		&	236.700		&	246.000			&	245.767			&	244.033	D			\\
128	&	18.667		&	17.833	B	&	10.367			&	12.933	O		&	243.367		&	247.000			&	247.000	d		&	245.833				\\ \hline
\multicolumn{9}{|c|}{ZDT3}																															\\ \hline
8	&	163.767		&	83.367		&	119.467			&	Equiv-D			&	176.533		&	226.400			&	207.700			&	Equiv-D				\\
16	&	108.467		&	49.433		&	47.700	D		&	Equiv-O			&	205.333		&	238.467			&	232.933			&	Equiv-O				\\
32	&	72.133		&	44.333		&	31.833			&	36.533	O		&	225.100		&	243.900			&	242.433			&	242.000	O			\\
64	&	54.867		&	57.600	B	&	40.633			&	29.300			&	236.333		&	245.933			&	245.733			&	244.000	D			\\
128	&	50.600		&	72.133		&	49.033	B		&	34.933			&	243.200		&	247.000			&	247.000	D		&	245.833				\\ \hline
\multicolumn{9}{|c|}{ZDT6}																															\\ \hline
8	&	22.133		&	13.833		&	14.433	D		&	Equiv-D			&	175.733		&	226.100			&	207.767			&	Equiv-D				\\
16	&	20.767		&	15.867		&	13.300	D		&	Equiv-O			&	205.033		&	238.433			&	232.933			&	Equiv-O				\\
32	&	22.600		&	10.700		&	10.033	D		&	11.967	DO		&	225.833		&	243.667			&	242.500			&	242.000	O			\\
64	&	19.033		&	11.267		&	10.533	D		&	10.567	DO		&	236.433		&	245.900			&	245.833			&	244.000	D			\\
128	&	15.800		&	24.500		&	11.600			&	12.233	DO		&	243.800		&	247.000			&	247.000	D		&	245.533				\\ \hline
\end{tabular}
}
\caption{Average number of generations and average number of solutions per front, obtained after 30 runs per configuration, for the 4 methods compared: baseline (B), disjoint (D), overlapped (O) and \textcolor{blue}{automatic} overlapped (A), using a chromosome length of 2048 dimensions. Acronyms next to values indicate that there is not significant difference with respect to that method for that value. Equiv-$X$ implies that value is the same as executing $X$, since both configurations would be the same.}
\label{tab:sols2048}
\end{table*}

@@

However, with 2048 dimensions, results show that dividing the chromosome produces an improvement in all the quality indicators using the automatic ($A$) version (Table \ref{tab:results2048}), even improving the overlapping and disjoint methods. Therefore, there exist some kind of limit point in chromosome length where one method will be preferable to another, besides depending on the number of islands and population size.

\begin{figure}
\centering
\includegraphics[width=12cm]{plot_zdt1_512.png}
\caption{\textcolor{blue}{PFs obtained in one arbitrary run for the ZDT1 problem with a chromosome size ($L$) of 512 for each combination of number of islands ($P$) and method: baseline ($B$), disjoint ($D$), overlapped ($O$) and automatic overlapped ($A$). Results for 8 and 16 islands of $A$ are equivalent of running $D$ and $O$ respectively. }}
\label{fig:plot_zdt1_512}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=12cm]{plot_zdt2_512.png}
\caption{\textcolor{blue}{PFs obtained in one arbitrary run for the ZDT2 problem with a chromosome size ($L$) of 512 for each combination of number of islands ($P$) and method: baseline ($B$), disjoint ($D$), overlapped ($O$) and automatic overlapped ($A$). Results for 8 and 16 islands of $A$ are equivalent of running $D$ and $O$ respectively. }}
\label{fig:plot_zdt2_512}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=12cm]{plot_zdt3_512.png}
\caption{\textcolor{blue}{PFs obtained in one arbitrary run for the ZDT3 problem with a chromosome size ($L$) of 512 for each combination of number of islands ($P$) and method: baseline ($B$), disjoint ($D$), overlapped ($O$) and automatic overlapped ($A$). Results for 8 and 16 islands of $A$ are equivalent of running $D$ and $O$ respectively. }}
\label{fig:plot_zdt3_512}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=12cm]{plot_zdt6_512.png}
\caption{\textcolor{blue}{PFs obtained in one arbitrary run for the ZDT6 problem with a chromosome size ($L$) of 512 for each combination of number of islands ($P$) and method: baseline ($B$), disjoint ($D$), overlapped ($O$) and automatic overlapped ($A$). Results for 8 and 16 islands of $A$ are equivalent of running $D$ and $O$ respectively. }}
\label{fig:plot_zdt6_512}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=12cm]{plot_zdt1_2048.png}
\caption{\textcolor{blue}{PFs obtained in one arbitrary run for the ZDT1 problem with a chromosome size ($L$) of 2048 for each combination of number of islands ($P$) and method: baseline ($B$), disjoint ($D$), overlapped ($O$) and automatic overlapped ($A$). Results for 8 and 16 islands of $A$ are equivalent of running $D$ and $O$ respectively. }}
\label{fig:plot_zdt1_2048}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=12cm]{plot_zdt2_2048.png}
\caption{\textcolor{blue}{PFs obtained in one arbitrary run for the ZDT2 problem with a chromosome size ($L$) of 2048 for each combination of number of islands ($P$) and method: baseline ($B$), disjoint ($D$), overlapped ($O$) and automatic overlapped ($A$). Results for 8 and 16 islands of $A$ are equivalent of running $D$ and $O$ respectively. }}
\label{fig:plot_zdt2_2048}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=12cm]{plot_zdt3_2048.png}
\caption{\textcolor{blue}{PFs obtained in one arbitrary run for the ZDT3 problem with a chromosome size ($L$) of 2048 for each combination of number of islands ($P$) and method: baseline ($B$), disjoint ($D$), overlapped ($O$) and automatic overlapped ($A$). Results for 8 and 16 islands of $A$ are equivalent of running $D$ and $O$ respectively. }}
\label{fig:plot_zdt3_2048}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=12cm]{plot_zdt6_2048.png}
\caption{\textcolor{blue}{PFs obtained in one arbitrary run for the ZDT6 problem with a chromosome size ($L$) of 2048 for each combination of number of islands ($P$) and method: baseline ($B$), disjoint ($D$), overlapped ($O$) and automatic overlapped ($A$). Results for 8 and 16 islands of $A$ are equivalent of running $D$ and $O$ respectively. }}
\label{fig:plot_zdt6_2048}
\end{figure}
